---
title: "R Notebook"
author: "Carina Huerta"
output:
  html_document:
    df_print: paged
---



#9 Introduction
- You will learn about tibbles, data import, tidy data, relational data, strings, factors, data and times

#10 Tibbles
  - Tibbles are data frames
### 10.1.1
```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
```
 - had to download magrittr in order to be able to use %>% 
## 10.2 Creating tibbles

```{r}
as_tibble(iris)
```


```{r}
tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)
```

- non-syntactic names - possible for a tibble to have column names that are not valid R variable names

```{r}
tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb
```


- tribble() is customised for data entry in code: column headings are defined by formulas and entries are separated by commas. 
```{r}
tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

## 10.3 Tibbles vs. data.frame

### 10.3.1 Printing

- Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on screen.


```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)

```

- Tibbles are designed so that you don’t overwhelm your console when you print large data frames.
    - First, you can explicitly print() the data frame and control the number of rows (n) and the width of the display. width = Inf will display all columns
```{r}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```
    
```{r}
nycflights13::flights %>% 
  View()
```

## 10.3.2 Subsetting

-  to pull out a single variable, you need some new tools, $ and [[. [[ can extract by name or position; $ only extracts by name

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

df$x
df[["x"]]
df[[1]]
```
- To use these in a pipe, you’ll need to use the special placeholder '.'


```{r}
df %>% .$x
df %>% .[["x"]]
```

## 10.4 Interacting with older code


```{r}
class(as.data.frame(tb))
```
- With base R data frames, [ sometimes returns a data frame, and sometimes returns a vector

## 10.5 Exercises

1).

- A tibble will print out the first ten observations
```{r}
mtcars
as_tibble(mtcars)
```


2).
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

- The $ operator will match any column name that starts with the name following it
- With data.frames, with [ the type of object that is returned differs on the number of columns

3).
If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?
in order to extract the variable you would use the double bracket
    - like df[[var]] 
cannot use the dollar sign because df$var would look for a column named var

4).
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

 extrat  ariable 1
```{r}
annoying[["1"]]
```
  create a scatter plot of 1 vs. 2
```{r}
ggplot(annoying, aes(x = `1`, y = `2`)) +
  geom_point()
```
  add a new column 3 which is 2 divided by 1
```{r}
annoying[["3"]] <- annoying[["2"]] / annoying[["1"]]
```
  rename the columns to one, two, and three, run
```{r eval=FALSE}
annoying <- rename(annoying, one = `1`, two = `2`, three = `3`)
glimpse(annoying)
```
  
5).

  - The function tibble::enframe() converts named vectors to a data frame with names and values

6).

  
- x	- Object to show

- n	- Number of rows to show. If NULL, the default, will print all rows if less than option tibble.print_max. Otherwise, will print tibble.print_min rows

- width - Width of text output to generate. This defaults to NULL, which means use getOption("tibble.width") or (if also NULL) getOption("width"); the latter displays only the columns that fit on one screen. You can also set options(tibble.width = Inf) to override this default and always print all columns

- n_extra	- Number of extra columns to print abbreviated information for, if the width is too small for the entire tibble. If NULL, the default, will print information about at most tibble.max_extra_cols extra columns


# 11 Data import

## 11.2 Getting started

- read_csv() reads comma delimited files, read_csv2() reads semicolon separated files (common in countries where , is used as the decimal place), read_tsv() reads tab delimited files, and read_delim() reads in files with any delimiter.

- read_fwf() reads fixed width files. You can specify fields either by their widths with fwf_widths() or their position with fwf_positions()
  - read_table() reads a common variation of fixed width files where columns are separated by white space.

- read_log() reads Apache style log files
  - (But also check out webreadr which is built on top of read_log() and provides many more helpful tools.)


```{r eval=FALSE}
heights<-read_csv("data/heights.csv")
read_csv("a,b,c
1,2,3
4,5,6")

```
```{r eval=FALSE}
read_csv("The first line of metadata
  The second line of metadata
  x,y,z
  1,2,3", skip = 2)

```

 ### 11.2.1 Exercises
 
 1). 
 
 - read_delim() function with the argument delim="|"
```{r eval=FALSE}
read_delim(file, delim = "|")
```
 
2).

```{r eval=FALSE}
union(names(formals(read_csv)), names(formals(read_tsv)))
```
   

3). 

- read_fwf() which reads “fixed-width formats”, is col_positions which tells the function where data columns begin and end

5).

```{r}
read_csv("a,b\n1,2,3\n4,5,6")
```
two columns are specified in the header “a” and “b”
the rows have three columns, so the last column is dropped.


```{r}
read_csv("a,b,c\n1,2\n1,2,3,4")
```
numbers of columns in the data do not match the number of columns in the header (three)
row one, there are only two values, so column c is set to missing
row two, there is an extra value, and that value is dropped.

```{r}
read_csv("a,b\n\"1")
```
opening quote "1 is dropped because it is not closed, and a is treated as an integer.
```{r}
read_csv("a,b\n1,2\na,b")
```
“a” and “b” are treated as character vectors since they contain non-numeric strings.
```{r}
read_csv("a;b\n1;3")
```
values are separated by “;” instead of “,”. 

## 11.3 Parsing a vector

```{r eval=FALSE}
str(parse_logical(c("TRUE", "FALSE", "NA")))
str(parse_integer(c("1", "2", "3")))
str(parse_date(c("2010-01-01", "1979-10-14")))
parse_integer(c("1", "231", ".", "456"), na = ".")
x <- parse_integer(c("123", "345", "abc", "123.45"))

```

- parse_logical() and parse_integer() parse logicals and integers respectively. There’s basically nothing that can go wrong with these parsers so I won’t describe them here further.

- parse_double() is a strict numeric parser, and parse_number() is a flexible numeric parser. These are more complicated than you might expect because different parts of the world write numbers in different ways.

- parse_character() seems so simple that it shouldn’t be necessary. But one complication makes it quite important: character encodings.

- parse_factor() create factors, the data structure that R uses to represent categorical variables with fixed and known values.

- parse_datetime(), parse_date(), and parse_time() allow you to parse various date & time specifications. These are the most complicated because there are so many different ways of writing dates.

### 11.3.1 Exercises

1). 

- date and time formats: date_names, date_format, and time_format
- time zone: tz
- numbers: decimal_mark, grouping_mark
- encoding: encoding
2).

decimal_mark is set to the comma ",", then the grouping mark is set to the period "."
If the grouping mark is set to a period, then the decimal mark is set to a comma

3). 
```{r eval=FALSE}
au_locale <- locale(date_format = "%d/%m/%Y")
parse_date("02/01/2006", locale = au_locale)

```



5).

delimiter. 
The function read_csv() uses a comma, while read_csv2() uses a semi-colon (;)
Using a semi-colon is useful when commas are used as the decimal point (as in Europe).
6).

Programs that identify the encoding of text include:

- readr::guess_encoding()
- stringi::str_enc_detect()
- iconv
- chardet (Python)

7).

```{r eval=FALSE}
parse_date(d1, "%B %d, %Y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3, "%d-%b-%Y")
parse_date(d4, "%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1, "%H%M")

```


## 11.4 Parsing a file

```{r eval=FALSE}
guess_parser("2010-10-01")
guess_parser("15:01")
guess_parser(c("TRUE", "FALSE"))
guess_parser(c("1", "5", "9"))
guess_parser(c("12,352,561"))
str(parse_guess("2010-10-10"))
```

- logical: contains only “F”, “T”, “FALSE”, or “TRUE”.
- integer: contains only numeric characters (and -).
- double: contains only valid doubles (including numbers like 4.5e-5).
- number: contains valid doubles with the grouping mark inside.
- time: matches the default time_format.
- date: matches the default date_format.
- date-time: any ISO8601 date.




## 11.5 Writing to a file
```{r eval=FALSE}
write_csv(challenge, "challenge.csv")
```
- write_rds() and read_rds() are uniform wrappers around the base functions readRDS() and saveRDS()
```{r}
write_rds(challenge, "challenge.rds")
read_rds("challenge.rds")

```




## 11.6 Other types of data
- haven reads SPSS, Stata, and SAS files.

- readxl reads excel files (both .xls and .xlsx).

- DBI, along with a database specific backend allows you to run SQL queries against a database and return a data frame.














